---
title: "K-means clustering iris dataset"
author: "Germán Hernández"
date: "31/7/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Summary
This applications let us see the clustering made by the k-means technique, using the data set `iris`.
We can increase the number of cluster that we wants until 9, and choose between the variables:
- Sepal.Widht
- Sepal.Lenght
- Petal.Lenght and
- Petal.Width

k-means only works with numerical variables, so don't give the user the option to select a categorical variable.

## Reproducible Pitch

This is a task for the data science coursera course about developing data products for week 4. As part of this, I have created a shiny app and deployed it on the shiny server. The link is  https://germanhdzc.shinyapps.io/kmeans/ . The code can be found at https://github.com/GermanHdzC/Developing-Data-Products-App.

## Exploratory Data Analysis

First of all we show a summary of Iris dataset to see the variables that we have

```{r}
summary(iris)
```

## UI code

```{r cars, echo = TRUE, eval=FALSE}
library(ggplot2)
library(shiny)

# Define UI for application that draws a histogram
vars <- setdiff(names(iris), "Species")

ui = pageWithSidebar(
    headerPanel('Iris k-means clustering'),
    sidebarPanel(
        selectInput('xcol', 'X Variable', vars),
        selectInput('ycol', 'Y Variable', vars, selected = vars[[2]]),
        numericInput('clusters', 'Cluster count', 3, min = 1, max = 9)
    ),
    mainPanel(
        plotOutput('plot1')
    )
)
```

## Server code

```{r, echo=TRUE, eval=FALSE}
server = function(input, output, session) {
    
    # Combine the selected variables into a new data frame
    selectedData <- reactive({
        iris[, c(input$xcol, input$ycol)]
    })
    
    clusters <- reactive({
        kmeans(selectedData(), input$clusters)
    })
    
    output$plot1 <- renderPlot({
        palette(c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3",
                  "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))
        
        par(mar = c(5.1, 4.1, 0, 1))
        plot(selectedData(),
             col = clusters()$cluster,
             pch = 20, cex = 3)
        points(clusters()$centers, pch = 4, cex = 4, lwd = 4)
    })
    
}
```


